### 2024-02-04 about.qmd
wip

---
title: "About"
---

I am Professor of Cognitive Science in the [Department of Psychology, University of Sheffield](https://www.sheffield.ac.uk/psychology) and Senior Research Fellow at the [Research on Research Institute](https://researchonresearch.org/).

### Research Interests

- Decision Making, Bias
- Reasoning
- Metaresearch
- Data Science
- Experimental Design

### Education

::: {.d-flex .justify-content-between}
::: {}
**New York University**  
PhD in clevrness
Not really
:::
::: {.text-end}
New York, NY  
Granted May 2016
:::
:::
- Some texthere

### Mastodon

<iframe allowfullscreen sandbox="allow-top-navigation allow-scripts allow-popups allow-popups-to-escape-sandbox" src="https://mastofeed.com/apiv2/feed?userurl=https%3A%2F%2Fmastodon.online%2Fusers%2Ftomstafford&theme=light&size=100&header=true&replies=false&boosts=false"></iframe>