### 2024-02-04 index.qmd
wip

---
title: "Tom Stafford, Professor of Cognitive Science"
format: html
page-layout: full
image: /tom_2020.jpg
---

::: {.html}
<style>
  iframe {
    width: 100%;
    aspect-ratio: 500 / 800;
    max-width: 500px; /* Optional */
  }
</style>
:::

I study learning and decision making. Much of my research looks at risk and bias, and their management, in decision making. I am also interested in skill learning, using measures of behaviour informed by work done in computational theory and neuroscience. More recently a strand of my research looks at complex decisions, and the psychology of reason, argument and persuasion.

HI

## Updates on <mastodon.online/@tomstafford>

<iframe allowfullscreen sandbox="allow-top-navigation allow-scripts allow-popups allow-popups-to-escape-sandbox" src="https://mastofeed.com/apiv2/feed?userurl=https%3A%2F%2Fmastodon.online%2Fusers%2Ftomstafford&theme=light&size=100&header=true&replies=true&boosts=false"></iframe>
Colophon

Website made with [Quarto](https://quarto.org), hosted on [github](https://github.com/tomstafford/tomstafford.github.io). Styles taken from <https://ddimmery.com/posts/quarto-website/>