### 2024-02-04 about.qmd
intro text

---
title: "About"
---

::: {.column-page-inset-left}
I study learning and decision making. Much of my research looks at risk and bias, and their management, in decision making. I am also interested in skill learning, using measures of behaviour informed by work done in computational theory and neuroscience. More recently a strand of my research looks at complex decisions, and the psychology of reason, argument and persuasion.
:::

### Research Interests

- Decision Making, Bias
- Reasoning
- Metaresearch
- Data Science
- Experimental Design