### 2024-05-12 publications.qmd
update

---
title: "Publications"
---

[Google scholar profile](https://scholar.google.co.uk/citations?hl=en&user=bMEna2UAAAAJ&view_op=list_works&sortby=pubdate), ORCID ID : [0000-0002-8089-9479](http://orcid.org/0000-0002-8089-9479)

# Theme highlights

--------------------

## Data intensive psychology:

-   Panagiotidi, M., Zavlis, O., Jones, M., & Stafford, T. (2023). [The three-dimensional community structure of ADHD traits captured by the ASRS scale: An Exploratory Graph Analysis](http://doi.org/10.1002/mpr.1997). International Journal of Methods in Psychiatric Research.  http://doi.org/10.1002/mpr.1997

-   Stafford, T., and Vaci, N. (2022). [Maximizing the potential of digital games for understanding skill acquisition](https://doi.org/10.1177/09637214211057841). Current Directions in Psychological Science, 31(1), 49-55.  <https://doi.org/10.1177/09637214211057841> [[web/mobile version](https://journals.sagepub.com/doi/full/10.1177/09637214211057841)]

-   Bannard, C., Leriche, M., Bandmann, O., Brown, C., Ferracane, E., Sánchez-Ferro, A., Obeso, J., Redgrave, P. and Stafford, T. (2019). [Reduced habit-driven errors in Parkinson's Disease](https://www.nature.com/articles/s41598-019-39294-z). Nature Scientific Reports 9(1), 3423. ([pre-print](https://psyarxiv.com/b25tf/))

## Metaresearch

-   Jones, A., Stafford, T., & Petrovskaya, E. (under review). [Exploring the multiverse of analysis options for the Addiction Stroop](https://psyarxiv.com/prfu2/).

-   Stafford T, Rombach I, Hind D et al. (2023) [Where next for partial randomisation of research funding? The feasibility of RCTs and alternatives](https://wellcomeopenresearch.org/articles/8-309/v1) Wellcome Open Res 2023, 8:309 (https://doi.org/10.12688/wellcomeopenres.19565.1)

-   Bendiscioli, Sandra; Firpo, Teo; Bravo-Biosca, Albert; Czibor, Eszter; Garfinkel, Michele; Stafford, Tom; et al. (2022): [The Experimental Research Funder's Handbook](https://rori.figshare.com/articles/report/The_experimental_research_funder_s_handbook_final_version_/19459328) (Revised edition, June 2022, ISBN 978-1-7397102-0-0). Research on Research Institute. Report. https://doi.org/10.6084/m9.figshare.19459328.v2

-   Silberzahn, R. et al (2018). [Many analysts, one dataset: Making transparent how variations in analytical choices affect results](http://journals.sagepub.com/doi/full/10.1177/2515245917747646). Advances in Methods and Practices in Psychological Science, 1(3), 337-356.

    -   See also : '[Science isn't broken](https://fivethirtyeight.com/features/science-isnt-broken/)' and the [interactive visualisation by Mat Evans which allowed us to find an important confound in the source data](https://github.com/mathewzilla/redcard/blob/master/Crowdstorming_visualisation.ipynb).

    -   Ploner, S., & Stafford, T. (2021). [How analysis strategy affects analysis results: Assessing results space and structure of Silberzahn et al. (2018) through model specification](https://osf.io/preprints/psyarxiv/nzskm). Retrieved from osf.io/preprints/psyarxiv/b2hm7

## Decision making & reasoning

-   Stafford, T., Pirrone, A., Croucher, M., and Krystalli, A. (2020). [Quantifying the benefits of using decision models with response time and accuracy data](https://link.springer.com/article/10.3758/s13428-020-01372-w). Behaviour Research Methods, 52, 2142--2155. [doi.org/10.3758/s13428-020-01372-w](https://doi.org/10.3758/s13428-020-01372-w) (pre-print, [web/mobile version](https://idiolect.org.uk/papers/stafford2020/))

-   Brand, C. O., Brady, D., & Stafford, T. (2023, June 27). [The Ideological Turing Test: a behavioural measure of open-mindedness and perspective-taking](https://psyarxiv.com/2e9wn). https://doi.org/10.31234/osf.io/2e9wn

-   Karadzhov, G., Stafford, T., and Vlachos, A. (2023). [DeliData: A dataset for deliberation in multi-party problem solving](https://dl.acm.org/doi/10.1145/3610056). Proceedings of the ACM on Human-Computer Interaction, CSCW. https://doi.org/10.1145/3610056 [[arXiv](https://arxiv.org/abs/2108.05271)]

-   Karadzhov, G., Stafford, T., and Vlachos, A. (2022). [What makes you change your mind? An empirical investigation in online group decision-making conversations](https://www.delibot.xyz/publications/). arXiv preprint [arXiv:2207.12035](https://arxiv.org/abs/2207.12035)

    -   -   See also: [delibot.xyz](https://www.delibot.xyz/)

-   de Kock, C., Stafford, T., & Vlachos, A. (2022). [How to disagree well: Investigating the dispute tactics used on Wikipedia](https://arxiv.org/abs/2212.08353). arXiv preprint arXiv:2212.08353.

# Chronological

## In press, and under review

-   Adinugroho, I., Stafford, T., & Bentall, R. (2023, November 1). [The correlation between conspiracy mentality and vaccine intentions is moderated by social events: Evidence from longitudinal data during COVID-19 pandemic in the UK](https://osf.io/preprints/psyarxiv/5wcnz/). Retrieved from osf.io/preprints/psyarxiv/5wcnz

-   Vardal, O., Karapanagiotidis, T., Stafford, T., Drachen, A. and Wade, A. (2023) [Unsupervised identification of internal perceptual states influencing psychomotor performance](https://www.biorxiv.org/content/10.1101/2023.09.08.556817v2). bioRxiv 2023.09.08.556817; doi: https://doi.org/10.1101/2023.09.08.556817

-   Copeland, A., Stafford, T., & Field, M. (2023, August 23). [Value-based decision-making in regular alcohol consumers following experimental manipulation of alcohol value](https://osf.io/wcks6/). https://doi.org/10.31219/osf.io/wcks6

-   Panagiotidi, M., Zavlis, O., Jones, M., & Stafford, T. (2023, July 14). [The three-dimensional community structure of ADHD traits captured by the ASRS scale: An Exploratory Graph Analysis](https://psyarxiv.com/sakn9). Retrieved from psyarxiv.com/sakn9

-   Brand, C. O., Brady, D., & Stafford, T. (2023, June 27). [The Ideological Turing Test: a behavioural measure of open-mindedness and perspective-taking](https://psyarxiv.com/2e9wn). https://doi.org/10.31234/osf.io/2e9wn

-   Zhu, J., Dommett, K., & Stafford, T. (2023). [What makes online political ads unacceptable? Interrogating public attitudes to inform regulatory responses](https://preprints.apsanet.org/engage/apsa/article-details/6462be37f2112b41e99fe9bf). APSA Preprints. doi:10.33774/apsa-2023-rxw1q This content is a preprint and has not been peer-reviewed.

-   Jones, A., Stafford, T., & Petrovskaya, E. (2022). [Exploring the multiverse of analysis options for the Addiction Stroop](https://psyarxiv.com/prfu2/).

-   Panagiotidi, M., Stafford, T., and Overton, P. (submitted). The relationship between cognitive control and task switching and problematic phone use. [Email for document]

-   Karadzhov, G., Stafford, T., and Vlachos, A. (2022). [What makes you change your mind? An empirical investigation in online group decision-making conversations](https://www.delibot.xyz/publications/). arXiv preprint [arXiv:2207.12035](https://arxiv.org/abs/2207.12035)

    -   -   See also: [delibot.xyz](https://www.delibot.xyz/)

-   de Kock, C., Stafford, T., & Vlachos, A. (2022). [How to disagree well: Investigating the dispute tactics used on Wikipedia](https://arxiv.org/abs/2212.08353). arXiv preprint arXiv:2212.08353.

## 2023

-   Karadzhov, G., Stafford, T., and Vlachos, A. (2023). [DeliData: A dataset for deliberation in multi-party problem solving](https://dl.acm.org/doi/10.1145/3610056). Proceedings of the ACM on Human-Computer Interaction, CSCW. https://doi.org/10.1145/3610056 [[arXiv](https://arxiv.org/abs/2108.05271)]

-   Stafford T, Rombach I, Hind D et al. (2023) [Where next for partial randomisation of research funding? The feasibility of RCTs and alternatives](https://wellcomeopenresearch.org/articles/8-309/v1) Wellcome Open Res 2023, 8:309 (https://doi.org/10.12688/wellcomeopenres.19565.1)

-   Farag, Y., Brand, C. O., Amidei, J., Piwek, P., Stafford, T., Stoyanchev, S., & Vlachos, A. (2023). Opening up Minds with Argumentative Dialogues. SiCON 2023 (co-located with ACL 2023) [[preprint]](https://arxiv.org/abs/2301.06400)

-   Dommett, K., Mensah, S. A., Zhu, J., Stafford, T., & Aletras, N. (2023). [Is There a Permanent Campaign for Online Political Advertising? Investigating Partisan and Non-Party Campaign Activity in the UK between 2018--2021](https://www.tandfonline.com/doi/full/10.1080/15377857.2023.2175102). Journal of Political Marketing, 1-19.

-   Varbanov V., Overton, P., and Stafford, T. (2023). [ADHD and ASD traits are indirectly associated with sensory changes through anxiety](https://link.springer.com/article/10.1007/s12144-022-04217-1) Curr Psychol (2023). <https://doi.org/10.1007/s12144-022-04217-1> [[preprint](https://psyarxiv.com/dchbs)].

## 2022

-   Vardal, O., Bonometti, V., Drachen, A., Wade, A., & Stafford, T. (2022). [Mind the gap: Distributed practice enhances performance in a MOBA game](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0275843). Plos one, 17(10), e0275843.

-   Brand, C. O., & Stafford, T. (2022). [Using dialogues to increase positive attitudes towards COVID-19 vaccines in a vaccine-hesitant UK population](https://royalsocietypublishing.org/doi/full/10.1098/rsos.220366). Royal Society Open Science, 9(10), 220366.

-   Copeland, A., Stafford, T., Acuff, S.F., Murphy, J.G. and Field, M. (in press). [Behavioral economic and value-based decision-making constructs that discriminate current heavy drinkers versus people who reduced their drinking without treatment](https://psyarxiv.com/5jhdg/). Psychology of Addictive Behaviors.

-   Bendiscioli, Sandra; Firpo, Teo; Bravo-Biosca, Albert; Czibor, Eszter; Garfinkel, Michele; Stafford, Tom; et al. (2022): [The Experimental Research Funder's Handbook](https://rori.figshare.com/articles/report/The_experimental_research_funder_s_handbook_final_version_/19459328) (Revised edition, June 2022, ISBN 978-1-7397102-0-0). Research on Research Institute. Report. https://doi.org/10.6084/m9.figshare.19459328.v2

-   Copeland et al (in press). Methodological issues with value-based decision-making (VBDM) tasks: the effect of trial wording on evidence accumulation outputs from the EZ drift-diffusion model. Cogent.

-   Stafford, T. (in press). [Evidence for the rationalisation phenomenon is exaggerated](https://psyarxiv.com/aygh6/). Behavioural and Brain Sciences (commentary)

-   Stafford, T., and Vaci, N. (2022). [Maximizing the potential of digital games for understanding skill acquisition](https://doi.org/10.1177/09637214211057841). Current Directions in Psychological Science, 31(1), 49-55.  <https://doi.org/10.1177/09637214211057841> [[web/mobile  version](https://journals.sagepub.com/doi/full/10.1177/09637214211057841)]

-   Pirrone, A., Reina, A, Stafford, T., Marshall, J.A.R., and Gobet, F. (2022). [Magnitude-sensitivity: Rethinking decision-making](https://www.sciencedirect.com/science/article/pii/S1364661321002618?casa_token=W5lKMgVueasAAAAA:GDwwQu4GBq4QtdbTWvjt7Y6-OY9kr7tgT0nGjQ1nu8a74laE5TCtmRotkpw5vQO6nQ8q14gH). Trends in Cognitive Sciences, 26(1), 66-80 [[OA version here](https://www.giovannireina.com/pdf/pirrone-TiCS2022.pdf)]

## 2021

-   Stafford, T. and Brand, C.O. (2021). [Commercial involvement in academic publishing is key to research reliability and should face greater public scrutiny](https://osf.io/preprints/metaarxiv/rjmvh). Evidence submitted to the House of Common Science and Technology Committee inquiry on reproducibility and research integrity.

-   Copeland, A., Stafford, T., and Field, M. (2021). Recovery from addiction: A synthesis of perspectives from behavioral economics, psychology, and decision modeling. In The Handbook of Alcohol Use (pp. 563-579). Academic Press.

-   Brand, C. O. and Stafford, T. (2021). Review 1: "Information Delivered by a Chatbot Has a Positive Impact on COVID-19 Vaccines Attitudes and Intentions." Rapid Reviews COVID-19. <https://doi.org/10.1162/2e3983f5.237d4808>

-   Ploner, S., & Stafford, T. (2021). [How analysis strategy affects analysis results: Assessing results space and structure of Silberzahn et al. (2018) through model specification](https://osf.io/preprints/psyarxiv/b2hm7). Retrieved from osf.io/preprints/psyarxiv/b2hm7

## 2020

-   Stafford, T., Pirrone, A., Croucher, M., and Krystalli, A. (2020). [Quantifying the benefits of using decision models with response time and accuracy data](https://link.springer.com/article/10.3758/s13428-020-01372-w). Behaviour Research Methods, 52, 2142--2155. [doi.org/10.3758/s13428-020-01372-w](https://doi.org/10.3758/s13428-020-01372-w) (pre-print, [web/mobile version](https://idiolect.org.uk/papers/stafford2020/))

    -   See also the interactive data explorer: sheffield-[university.shinyapps.io/decision_power](https://sheffield-university.shinyapps.io/decision_power/)

-   Dhir, N., Edman, M., Sanchez Ferro, A., Stafford, T., and Bannard, C. (2020). [Identifying robust markers of Parkinson's Disease in typing behaviour using a CNN-LSTM network](https://www.aclweb.org/anthology/2020.conll-1.47/). In Proceedings of Conference on Computational Natural Language Learning (CoNLL 2020).

-   Panagiotidi, M., Overton, P. and Stafford, T. (2020). [The relationship between sensory processing sensitivity and Attention Deficit Hyperactivity Disorder traits: a spectrum approach](https://www.sciencedirect.com/science/article/pii/S0165178120331383). Psychiatry Research, 293, 113477.

-   Scaife, R., Stafford, T., Bunge, A., and Holroyd, J. (2020). [To blame? The effects of moralized feedback on implicit racial bias](https://www.collabra.org/article/10.1525/collabra.251/). Collabra: Psychology, 6(1), 30. DOI: http://doi.org/10.1525/collabra.251

-   Pirrone, A., Illin, J., Stafford, T., Milne, E. (2020). [A diffusion model decomposition of orientation discrimination in children with Autism Spectrum Disorder](https://www.tandfonline.com/eprint/VIIZBcGgYpItKcCGw27I/full). European Journal of Developmental Psychology, 17, 2, 213-230.

-   Williams, L., Butler, J., Thirkettle, M., Stafford, T., Quinlivan, B., MacGovern, E., ... and Hutchinson, M. (2020). [Slowed luminance reaction times in cervical dystonia: disordered superior colliculus processin](https://onlinelibrary.wiley.com/doi/full/10.1002/mds.27975)g. Movement Disorders, 35(5), 877-880.

-   Field, M., Heather, N., Murphy, J. G., Stafford, T., Tucker, J. A., & Witkiewitz, K. (2020). [Recovery from addiction: Behavioral economics and value-based decision making](https://psyarxiv.com/rfjwu/). Psychology of Addictive Behaviors, 34(1), 182--193. https://doi.org/10.1037/adb0000518

-   Gyurkovics, M., Stafford, T., Levita, L. (2020). [Cognitive control across adolescence: Dynamic adjustments and mind-wandering](https://www.ncbi.nlm.nih.gov/pubmed/31599622). Journal of Experimental Psychology: General, 149(6), 1017--1031.

-   Tierney, W., Hardy III, J. H., Ebersole, C. R., Leavitt, K., Viganola, D., Clemente, E. G., ... & Hiring Decisions Forecasting Collaboration. (2020). [Creative destruction in science](https://www.sciencedirect.com/science/article/pii/S0749597820303678). Organizational Behavior and Human Decision Processes, 161, 291-309.

## 2019

-   Hampsey, E., Overton, P.G., Stafford, T. (2019). [Microsaccade rate as a measure of drug response](https://bop.unibe.ch/JEMR/article/view/JEMR.12.6.12). Journal of Eye Movement Research, 12(6). https://doi.org/10.16910/jemr.12.6.12   Open data/analysis scripts/etc: <https://osf.io/je8hf/>

-   Bannard, C., Leriche, M., Bandmann, O., Brown, C., Ferracane, E., Sánchez-Ferro, A., Obeso, J., Redgrave, P. and Stafford, T. (2019). [Reduced habit-driven errors in Parkinson's Disease](https://www.nature.com/articles/s41598-019-39294-z). Nature Scientific Reports 9(1), 3423. ([pre-print](https://psyarxiv.com/b25tf/))

-   Holroyd, J., Scaife, R., Stafford, T. (in press). [What is implicit bias](http://onlinelibrary.wiley.com/doi/10.1111/phc3.12437/full)? Philosophy Compass, 12(10), e12437.

-   Holroyd, J., Scaife, R., Stafford, T. (in press). [Responsibility for Implicit Bias](http://onlinelibrary.wiley.com/doi/10.1111/phc3.12410/abstract). Philosophy Compass, 12(3), e12410.

-   Thirkettle, M., Thyoka, M., Fernandes, N., Gopalan, P., Stafford, T. Offiah, A.C. (2019). [Internet-based measurement of visual assessment skill of trainee radiologists: developing a sensitive tool](https://www.birpublications.org/doi/abs/10.1259/bjr.20180958). British Journal of Radiology, 92(1097), 20180958.

## 2018

-   Silberzahn, R. et al (2018). [Many analysts, one dataset: Making transparent how variations in analytical choices affect results](http://journals.sagepub.com/doi/full/10.1177/2515245917747646). Advances in Methods and Practices in Psychological Science, 1(3), 337-356. See also : '[Science isn't broken](https://fivethirtyeight.com/features/science-isnt-broken/)' and the [interactive visualisation by Mat Evans which allowed us to find an important confound in the source data](https://github.com/mathewzilla/redcard/blob/master/Crowdstorming_visualisation.ipynb).

-   Stafford, T., Holroyd, J., and Scaife, R. (2018). [Confronting bias in judging: A framework for addressing psychological biases in decision making](https://psyarxiv.com/nzskm). https://doi.org/10.31234/osf.io/nzskm

    -   See also: [Judiciary.uk: Biases in decision making](https://www.judiciary.uk/wp-content/uploads/2018/02/stafford-biases-in-decision-making-winter-2017.pdf).

-   Stafford, T. (2018). [Female chess players outperform expectations when playing men](http://journals.sagepub.com/doi/10.1177/0956797617736887). Psychological Science, 29(3), 429-436. (Preprint).

-   Pirrone, A., Azab, H., Hayden, B. Y., Stafford, T., and Marshall, J. A. R. (2018). [Evidence for the speed--value trade-off: Human and monkey decision making is magnitude sensitive](http://eprints.whiterose.ac.uk/109297/). Decision, 5(2), 129-142.

-   Panagiotidi, M., Overton, P.G., Stafford, T. (2018). [The relationship between ADHD traits and sensory sensitivity in the general population](https://www.sciencedirect.com/science/article/pii/S0010440X17302365). Comprehensive Psychiatry, 80, 179-185.

-   Kalfaoğlu, Ç., Stafford, T. and Milne, E. (2018). [Frontal theta band oscillations predict error correction and post error slowing in typing](https://www.ncbi.nlm.nih.gov/pubmed/28447844). Journal of Experimental Psychology: Human Perception and Performance, 44(1), 69-88.

## 2017

-   Stafford, T., Devlin, S., Sifa, R., and Drachen, A. (2017). [Exploration and skill acquisition in a major online game](https://mindmodeling.org/cogsci2017/papers/0615/paper0615.pdf). In Proceedings of The 39th Annual Meeting of the Cognitive Science Society.

-   Panagiotidi, M., Overton, P.G., Stafford, T. (2017). [Multisensory integration and ADHD-like traits: Evidence for an abnormal temporal integration window in ADHD](http://www.sciencedirect.com/science/article/pii/S000169181630302X). Acta Psychologica, 181, 10--17.

-   Panagiotidi, M., Overton, P.G., Stafford, T. (2017). [Co-occurrence of ASD and ADHD traits in the general population](http://eprints.staffs.ac.uk/3487/). Journal of Attentional Disorders.

-   Panagiotidi, M., Overton, P.G., Stafford, T. (2017). [Increased microsaccade rate in individuals with ADHD traits](https://bop.unibe.ch/index.php/JEMR/article/view/2963). Journal of Eye Movement Research. 10, 1.

-   Pirrone, A., Dickinson, A., Gomez, R., Stafford, T. and Milne, E. (2017). [Understanding perceptual judgement in autism spectrum disorder using the drift diffusion model](http://eprints.whiterose.ac.uk/107355/1/Perceptual%20Judgment%20in%20ASD%20final%20copy.pdf). Neuropsychology, 31 (2), 173--180.

-   Pirrone, A., Marshall, J. A., and Stafford, T. (2017). [A drift diffusion model account of the semantic congruity effect in a classification paradigm](https://jnc.psychopen.eu/article/view/79). Journal of Numerical Cognition, 3(1), 77--96.

-   Stafford, T. and Haasnoot, E. (2017). [Testing sleep consolidation in skill learning: a field study using an online game](http://onlinelibrary.wiley.com/wol1/doi/10.1111/tops.12232/abstract). Topics in Cognitive Science. 9(2), 485--496. ([data + code](https://osf.io/fckq8/))

-   Panagiotidi, M., Overton, P.G., Stafford, T. (2017). [Attention Deficit Hyperactivity Disorder-like traits and distractibility in the visual periphery](http://pec.sagepub.com/content/early/2016/11/28/0301006616681313.abstract). Perception, 46 (6), 665--678.

## 2016

-   Bednark J., Reynolds J., Stafford T., Redgrave P. and Franz E. (2016). [Action experience and action discovery in medicated individuals with Parkinson's disease](http://journal.frontiersin.org/article/10.3389/fnhum.2016.00427). Frontiers in Human Neuroscience, 10, 427. DOI 10.3389/fnhum.2016.00427.

-   Bertram, C., and Stafford, T. (2016). [Improving training for sensory augmentation using the science of expertise](http://www.sciencedirect.com/science/article/pii/S0149763415300634). Neuroscience & Biobehavioral Reviews, 68, 234--244.

-   Lu, Y., Stafford, T., and Fox, C. (2016). [Maximum saliency bias in binocular fusion](http://www.tandfonline.com/doi/abs/10.1080/09540091.2016.1159181). Connection Science, 28(3),258-269.

## 2015

-   Thirkettle, M., Stafford, T., and Offiah, A. (2015). Internet Based Measurement of Visual Expertise in Radiological Skill. Perception, 44, 44-45.

## 2014

-   Stafford, T. and Dewar, M. (2014). [Tracing the trajectory of skill learning with a very large sample of online game players](http://pss.sagepub.com/content/25/2/511). Psychological Science, 25(2) 511--518. See also: [Data and analysis code](https://github.com/tomstafford/axongame).

-   Baldassarre, G., Stafford, T., Mirolli, M., Redgrave, P., Ryan, R. M., and Barto, A. (2014). [Intrinsic motivations and open-ended development in animals, humans, and robots: an overview](http://journal.frontiersin.org/Journal/10.3389/fpsyg.2014.00985/full). Frontiers in Psychology, 5(985). doi: 10.3389/fpsyg.2014.00985 (introduction to Special Topic we edited).

-   Kalfaoğlu, Ç & Stafford, T. (2014). [Performance breakdown effects dissociate from error detection effects in typing](http://www.tandfonline.com/eprint/neYR97ser5C3sFI8pcuT/full). The Quarterly Journal of Experimental Psychology, 67(3), 508--524.

-   Stafford, T. (2014). [The perspectival shift: how experiments on unconscious processing don't justify the claims made for them](http://journal.frontiersin.org/Journal/10.3389/fpsyg.2014.01067/full). Frontiers in Psychology, 5, 1067. doi:10.3389/fpsyg.2014.01067.

-   Stafford, T., Elgueta, H., Cameron, H. (2014).[ Students' engagement with a collaborative wiki tool predicts enhanced written exam performance](http://www.researchinlearningtechnology.net/index.php/rlt/article/view/22797). Research in Learning Technology, 22, 22797. doi:[10.3402/rlt.v22.22797](http://dx.doi.org/10.3402/rlt.v22.22797).

-   Pirrone, A., Stafford, T., and Marshall, J. A. R. (2014). [When natural selection should optimise speed-accuracy trade-offs](http://www.frontiersin.org/decision%20neuroscience/10.3389/fnins.2014.00073/full). Frontiers in Neuroscience, 8(73). doi: 10.3389/fnins.2014.00073.

-   Redgrave, P., Vautrelle, N., and Stafford, T. (2014). [Interpretive conundrums when practice doesn't always make perfect](http://www.ncbi.nlm.nih.gov/pubmed/24167063). Movement Disorders: Official Journal of the Movement Disorder Society, 29(1), 7-10. doi:10.1002/mds.25726.

-   Kerdegari, H., Kim, Y., Stafford, T., and Prescott, T. J. (2014, June). [Centralizing bias and the vibrotactile funnelling illusion on the forehead](https://eprints.whiterose.ac.uk/100435/1/Eurohaptic-final%20version-HK.pdf). In International Conference on Human Haptic Sensing and Touch Enabled Computer Applications (pp. 55-62). Springer, Berlin, Heidelberg.

## 2013

-   Bertram, C., Evans, M. H., Javaid, M., Stafford, T., and Prescott, T. (2013). Sensory augmentation with distal touch: the tactile helmet project. In Biomimetic and Biohybrid Systems (pp. 24-35). Springer Berlin Heidelberg.

-   Thirkettle, M., Walton, T., Redgrave, P., Gurney, K., and Stafford, T. (2013). [No learning where to go without first knowing where you're coming from: action discovery is trajectory, not endpoint based](http://www.frontiersin.org/cognitive_science/10.3389/fpsyg.2013.00638/abstract). Frontiers in Cognitive Science, 4:, 638. doi:10.3389/fpsyg.2013.00638.

-   Walton, T., Thirkettle, M., Redgrave, P., Gurney, K. N., and Stafford, T. (2013). [The discovery of novel actions is affected by very brief reinforcement delays and reinforcement modality](http://www.tandfonline.com/doi/full/10.1080/00222895.2013.806108). Journal of Motor Behavior, 45(4), 351--360.

-   Bednark, J. G., Reynolds, J. N. J., Stafford, T., Redgrave, P., and Franz, E. A. (2013). [Creating a movement heuristic for voluntary action: Electrophysiological correlates of movement-outcome learning](http://www.sciencedirect.com/science/article/pii/S0010945211003224). Cortex, 49(3), 771-780. doi:10.1016/j.cortex.2011.12.005.

-   Thirkettle, M., Walton, T., Shah, A., Gurney, K., Redgrave, P., and Stafford, T. (2013). [The path to learning: Action acquisition is impaired when visual reinforcement signals must first access cortex](http://www.sciencedirect.com/science/article/pii/S0166432813000429). Behavioural Brain Research, 243, 267--272. doi:10.1016/j.bbr.2013.01.023.

## 2012

-   Stafford, T., Thirkettle, M., Walton, T., Vautrelle, N., Hetherington, L., Port, M., Gurney, K.N., Redgrave, P. (2012), [A novel task for the investigation of action acquisition](http://www.plosone.org/article/info%3Adoi%2F10.1371%2Fjournal.pone.0037749), PLoS One, 7(6), e37749.

-   Stafford, T. and Grimes, A. (2012). [Memory enhances the mere exposure effect](https://drive.google.com/file/d/1XuHvFVk3rbWHmuEHclPxFGSXSi8_Rp88/view?usp=sharing). (PDF, 626KB) Psychology & Marketing, 29, 12, 995--1003.

-   Stafford, T., and Bell, V. (2012). [Brain network: social media and the cognitive scientist](https://drive.google.com/file/d/1ouB70muf-9-XN_7RbeN86QQIU1tEn8lZ/view?usp=sharing). (PDF, 108KB) Trends in Cognitive Sciences, 16(10), 489--490. doi:10.1016/j.tics.2012.08.001.

## 2011

-   Stafford, T., & Gurney, K. N. (2011). [Additive factors do not imply discrete processing stages: a worked example using models of the Stroop Task](http://www.frontiersin.org/Cognitive_Science/10.3389/fpsyg.2011.00287/abstract). Frontiers in Psychology, 2. doi:10.3389/fpsyg.2011.00287.

-   Stafford, T., Ingram, L. and Gurney, K.N. (2011), [Pieron's Law holds during Stroop conflict: insights into the architecture of decision making](https://drive.google.com/file/d/19v6Bi6x01YBs5InJj8ALNNfhUP1ZrV6H/view?usp=sharing). (PDF, 374KB)  Cognitive Science 35, 1553--1566.

-   Yates, D.J. and Stafford, T. (2011), [Insights into the function and mechanism of saccadic decision making from targets scaled by an estimate of the cortical magnification factor](http://www.abrg.group.shef.ac.uk/!DATA/attachment/0102.Yates2011a.pdf), Cognitive Computation, 3, 89--93.

## Earlier key publications

-   Stafford, T. (2010). [How do we use computational models of cognitive processes?](https://drive.google.com/file/d/1U3JZPaUb_J-L-tj9t8CpevsvUTAcOyvr/view?usp=sharing) (PDF, 118KB) In Connectionist Models Of Neurocognition And Emergent Behaviour: From Theory to Applications. Proceedings of the 12th Neural Computation and Psychology Workshop, Birkbeck, University of London, 8-10 April 2010. World Scientific (pp. 326--342).

-   Stafford, T. (2009), [What use are computational models of cognitive processes?](https://drive.google.com/file/d/1KTWXcmsH9TWwVVagAfysf0ch-IQjzIDW/view?usp=sharing) (PDF, 56KB) In Mayor, J., Ruh, N.,  Plunkett, K. Connectionist Models of Behaviour and Cognition II: Proceedings of the 11th Neural Computation and Psychology Workshop. World Scientific.

-   Eiser, J. R., Stafford, T., Henneberry, J., and Catney, P. (2009). "Trust me, I'm a Scientist (Not a Developer)": [Perceived expertise and motives as predictors of trust in assessment of risk from contaminated land](https://drive.google.com/file/d/1mBtjfzW89lcLCWogK9ZYhIm3EJNFqCJq/view?usp=sharing). (PDF, 260KB)  Risk Analysis, 29(2), 288--297.

    -   see also [Why don't we trust the experts? (2016)](https://tomstafford.sites.sheffield.ac.uk/news/why-dont-we-trust-the-experts)

-   Stafford, T. and Gurney, K.N. (2007), [Biologically constrained action selection improves cognitive control in a model of the Stroop task](http://www.abrg.group.shef.ac.uk/!DATA/attachment/0046.stafford07.pdf), (External PDF) Philosophical Transactions of the Royal Society B: Biological Sciences, 362 (1485), 1671--1684.

-   Stafford, T. and Wilson, S. P. (2007), [Self-organisation can generate the discontinuities in the somatosensory map](https://drive.google.com/file/d/1IblZNhCxsEbgnL4K5x25qahwsXT5so4G/view?usp=sharing). (PDF, 347KB)  Neurocomputing, 70(10-12), 1932--1937.

-   Stafford, T. and Gurney, K. N. (2004), [The role of response mechanisms in determining reaction time performance: Pieron's Law revisited](https://drive.google.com/file/d/1173WABpGb9jbDOwU-i99X9zrU5szzlmL/view?usp=sharing). (PDF, 231KB)  Psychonomic Bulletin & Review, 11, 975--987.

-   Eiser, J. R., Fazio, R. H., Stafford, T. and Prescott, T. J. (2003), [Connectionist simulation of attitude learning: Asymmetries in the acquisition of positive and negative evaluations](https://drive.google.com/file/d/1bULXLeeUMv3wgq7g47JgLAZYfD_0ailg/view?usp=sharing). (PDF, 505KB)  Personality and Social Psychology Bulletin, 29, 1221--1235.