### 2024-05-12 done.qmd
update

---
title: "Done"
---

Recording things i've done which don't go under  [Papers](papers.html), 
[Talks](https://tomstafford.sites.sheffield.ac.uk/outputs/talks) or Newsletter: <https://tomstafford.substack.com/>. For current preoccupations, see [here](news.html).

## 2023

December 2023 : joined the PsyArXiv Scientific Advisory Board and approved as chair of the public engagement subcommittee. <https://osf.io/preprints/psyarxiv>

2023-11-10; Worked with Lingyu-Meng on a novel visualisation of how group discussion shifts people's opinions : <a href="Dynamics of group discussion in agreement-error space">https://lingyu-meng.github.io/convergence-valence_space/</a>

2023-11-07-: elected Chair of UKRN Institutional Leads group

Sept 2022 - Sept 2023 Coordinator for UKRN [Open Research programme](https://www.ukrn.org/open-research-programme/), [workstream 1: training](https://www.ukrn.org/ws1-training/)

2023 August made <https://tomstafford.github.io/imdb/>, showing change in crowdsourced ratings of summer movies:

<img src="https://tomstafford.github.io/imdb/plots/average_ratings_ordered.png" height=400>

2023 July Consulted on [University of Sheffield Guidelines on Generative AI](https://sheffield.libguides.com/genai/)

2023-05-23 Analysis of Open Editors data: <https://tomstafford.github.io/editors/>

<img src="https://tomstafford.github.io/editors/figs/gpa_vs_eds.png" height = 400>

2023-05-18 Chaired UKRN/RoRI panel on [The Future of Peer Review](https://www.ukrn.org/event/future-of-peer-review-may2023/). ([Recording](https://youtu.be/BMeSh6esehE), [Slides](https://osf.io/ce3ts/))

2023-05-05 Co-author on UKRN working paper on [Open Research Training Resources and Priorities](https://osf.io/s2f6k/), ([news item](https://www.ukrn.org/2023/05/05/ukrn-first-working-paper-on-training-priorities-and-resources/))

2023-05-10 Cited in Science, Innovation and Technology Committee report on [Reproducibility and research integrity](https://publications.parliament.uk/pa/cm5803/cmselect/cmsctech/101/summary.html). Written evidence:[Commercial involvement in academic publishing is key to research reliability and should face greater public scrutiny](https://osf.io/preprints/metaarxiv/rjmvh)

2023-04-15 Chaired [Straight to the Facts: Can we save public debate? A dialogue with Will Moy](https://www.eventbrite.co.uk/e/straight-to-the-facts-can-we-save-public-debate-a-dialogue-with-will-moy-tickets-546336496167)

2023-02-19 Updated my minimal working example of RMarkdown + papaja + tenzing for APA formatted manuscript preparation, to include a rights retention statement : <https://github.com/tomstafford/rmarkdown_apa>

2023-02-06 Teaching my MSc course in Data Management and Visualisation. Materials, all CC-BY, and containing many links and resources on all things research project management, data wrangling and #dataviz  here <https://tomstafford.github.io/psy6422/>

2023-02-01 and forever: taking part in [UCU strike action](https://mastodon.online/@tomstafford/109788334770383538)

2023-01-24 Attended launch of [N8 Research Partnership stands up for researchers with new Rights Retention statement](https://www.n8research.org.uk/n8-research-partnership-rights-retention-statement/)

2023-01-17 Celebrating [University of Sheffield announcing rights retention policy](https://www.researchprofessional.com/0/rr/news/uk/universities/2023/1/University-of-Sheffield-approves-policies-to-bolster-open-research.html)

2023-01-16 Retired the [Choice Engine](https://tomstafford.github.io/choice-engine-text/) bot

2023-01-09 Graduation for Psychological Research Method with Data Science MSc [class of 2021](https://mastodon.online/@tomstafford/109659193946308768)

2023-01-06 Updated: Experiment Design Checklist <https://tomstafford.github.io/psy-checklist/>

2023-01-05 [Christmas message 2022](https://tomstafford.sites.sheffield.ac.uk/news/christmas-message-2022)

2023-01-01 [Archive of tweets 2008-2022](https://idiolect.org.uk/tweets/)

# 2021

2021: [The Experiment Design Checklist](https://tomstafford.github.io/psy-checklist/)

# 2020

Led development and adoption of the [University of Sheffield statement on Open Research](https://www.sheffield.ac.uk/openresearch/university-statement-open-research)

2020: [The Letters of Pope Gregory VII - pilot project](https://figshare.com/articles/dataset/The_Letters_of_Pope_Gregory_VII_-_pilot_project/12781049) (with Charles West and George Litchfield)

# 2019

[Animated Itinerary of King John,](https://sites.google.com/d/1nQb4g6jqk7M0gwaD4Bvw7QKtBg0nkQRV/p/120fMmXEkjhEwEnsgf8fx5v1XKQrf3MdZ/edit)  [Mapping The Itinerary of King Edward I](https://figshare.com/articles/dataset/Mapping_The_Itinerary_of_King_Edward_I/8948699), [Mapping Papal Letters](https://tomstafford.github.io/PapalCorrespondence/) (with Charles West)

# 2018

2018: [@ChoiceEngine - an interactive essay about the psychology of free will](https://github.com/tomstafford/choice-engine-text)

# Earlier

2017 Organised [Mind Matters](https://tomstafford.sites.sheffield.ac.uk/sub-sites/mind-matters) 'a series of events based on various psychological research to bring knowledge to the general public and showcase Sheffield as a hub for psychology!'

2012 - 2017 Wrote a column for [BBC Future](https://www.bbc.com/future/columns/neurohacks)

2012-09 I wrote some of the code in LongArm, [a robot which draws labrynthine patterns](https://www.youtube.com/watch?v=lSuvSsDrEAw), by Matthias Jones

2012: [The Cognitive Science Safari](https://sites.google.com/d/1nQb4g6jqk7M0gwaD4Bvw7QKtBg0nkQRV/p/1Q0M0b1kRUveC9w0Li7Je0A1f1E8PzzYB/edit) was a street-tour of classic social psychology experiments, in Berlin as part of the BMW Guggenheim Lab.

2011: [The Tea Taste Test](https://sites.google.com/d/1nQb4g6jqk7M0gwaD4Bvw7QKtBg0nkQRV/p/1O7uZnZPiKBL7C6JiR0DLSgGCqVw86uYX/edit) -- explaining why statistics are important to psychology, using a controversy of how to make a cup of tea.