### 2024-05-12 supervision.qmd
update

---
title: "Supervision"
---

I am always enthusiastic to hear from potential PhD students. Please get in touch.

A PhD with me will involve using [robust](https://www.nature.com/articles/s41562-016-0021) and [open science](https://en.wikipedia.org/wiki/Open_science) methods to address theoretical ideas in cognitive science. Plus extensive mentoring on all aspects of the scholarly life, conducted in [Sheffield's best coffee shops](https://www.ourfaveplaces.co.uk/our-guides/top-10-coffee-shops/).

Specific projects I am currently seeking students for are [here](https://docs.google.com/document/d/1v9akVLjoYlR01V7SvGjYxUsDP2zCCbdvSXYB1NqoqcQ/edit?usp=sharing). Some advice about applying is [here](https://github.com/tomstafford/onboarding/).

# Current PhD and postdoc supervision

## Postdoc

-   [Amber Copeland](https://scholar.google.com/citations?user=JnNAldsAAAAJ&hl=en) (2018--2025) on "Reinforcer-specific value-based decision-making in persistence of and recovery from alcohol use disorder"

-   [Georgi Karadzhov](https://gkaradzhov.com/) (2024--2025) -- "DEliBots -- Deliberation Enhancing Bots" (at University of Cambridge)

## PhD

-   Indro Adinugroho (2021--2025) -- Investigating the Relation between Fundamentalism, Belief in Conspiracy Theories and Perceived Political Leadership toward Political Trust

-   Hazel Gordon (2021--2025) "[Informing citizens? Effects of imprints on digital political advertising](https://tomstafford.sites.sheffield.ac.uk/news/phd-opportunity-informing-citizens)"

-   Ioanna Spentza (2017--) -- Factors affecting Everyday Problem-Solving skills in Older Adults

### Second supervisor to...

-   Vesko Varbanov (2020-) ' Sensory processing in Autism Spectrum Conditions and Attention Deficit Hyperactivity Disorder'

# Previous supervision and examining

## As primary PhD supervisor

-   [Maria Panagiotidi](https://scholar.google.co.uk/citations?user=XcFf_6EAAAAJ&hl=en&oi=ao), PhD 2017 '[The Role of the Superior Colliculus in Attention Deficit Hyperactivity Disorder](https://etheses.whiterose.ac.uk/12124/)'

-   [Angelo Pirrone](https://scholar.google.co.uk/citations?user=UZzVps4AAAAJ&hl=en), PhD 2016 '[Decision modelling insights in cognition and adaptive decision making](http://etheses.whiterose.ac.uk/15841/)'

-   Despina Panagiotidou (co-supervised with Ana Vivas, City College, Thessaloniki), PhD 2015, 'Distractor, Temporal, and Inhibitory Variations of the Distractor Devaluation Effect of Visual Selective Attention'

-   [Tom Walton](http://www.abrg.group.shef.ac.uk/people/walton/), PhD 2012, '[The Discovery of Novel Actions](http://etheses.whiterose.ac.uk/2204/)'

-   David Yates, PhD 2012, 'Using eye-movements to elucidate architectures of object recognition and selection'

-   Cigir Kalfaoglu, PhD 2012, '[Errors and error-detection during skilled typing](http://etheses.whiterose.ac.uk/2924/)'

## As a secondary PhD supervisor

-   [Georgi Karadzhov](https://gkaradzhov.com/) (2019--2022) -- "DEliBots -- Deliberation Enhancing Bots" (at University of Cambridge)

-   Ozan Vardal (2018--2022) -- [Leveraging Big Data from eSports Games to Understand Human Skill Acquisition](http://iggi.org.uk/students/2018/ozan-vardal/) (At University of York)

-   [Elena Petrovskaya](https://www.elenapetrovskaya.com/) (2019--2023) -- Decision-making & e-sports (At University of York)

-   [Amber Copeland](https://scholar.google.com/citations?user=JnNAldsAAAAJ&hl=en) (2018--2022)  -- 'Value-based decision-making and self-control in alcohol-related behaviour change and recovery from addiction'

-   August Lindemer (2018-2022) 'Planetary Fever -- The Climate Change Communication of Health Professionals.'

-   Siham Albesisi (2018 - 2023) 'emotion regulation on ADHD children'

-   Stephanie Powell ' The relationship and effects of young children's use of technology'

-   [Máté Gyurkovics](https://scholar.google.hu/citations?user=XGk9OnEAAAAJ&hl=hu) --  'Affective factors in cognitive control: Behavioural and neural investigations in adults and adolescents'

-   Stephanie Armstrong 'Relation between the Executive Function development and the reading acquisition, from a neuropsychological view'

-   Andreas Bunge (Department of Philosophy, University of Sheffield), "The mental structure of prejudice", February 2018

-   Luana Nunes, "Ethological Decision Making with Non-Stationary Inputs Using MSPRT Based Mechanisms" May 2017

-   Alvin Pastore (Department of Computer Science), "Neuroeconomics, Reinforcement learning and the equity premium puzzle" January 2017

-   Stephanie Dunn (with Elizabeth Milne and Megan Freeth), PhD 2016 'Electrophysiological correlates of visual search in the autism spectrum'

-   [Hamideh Kerdegari](http://scholar.google.co.uk/citations?user=BZdPw4MAAAAJ&hl=en) (supervised with Tony Prescott), PhD 2016 'Head-mounted Sensory Augmentation System for Navigation in Low Visibility Environments'

-   Tom Bullock (supervised with Elizabeth Milne and Dick Eiser), 2012, 'Crossmodal Load and Selective Attention'

## Research Associate Supervision

-   Zarha Arjmandi Lari, 2022 - 2023

-   Dagmar Adamcová, 2017--2018: "Quit while you're ahead: a surprising interaction between game performance and motivation" ([Blog post](https://tomstafford.sites.sheffield.ac.uk/news/quit-while-youre-ahead), [Jupyter notebook](https://github.com/dagmaradamcova/axon-drop-out/blob/master/axon_drop_out.ipynb)). Erasmus exchange

## Post doctoral supervision

-   [Junyan Zhu](https://www.sheffield.ac.uk/politics/people/academic-staff/junyan-zhu) (2020--2023)

-   [Lotty Brand](https://lottybrand.wordpress.com/) (2020--2023)

-   Mariana Leriche, 2014--2016

-   Robin Scaife, 2013--2017

-   Jen Lewis, 2012--2013

-   Craig Bertram, 2012--2013

-   [Martin Thirkettle](https://www.shu.ac.uk/about-us/our-people/staff-profiles/martin-thirkettle), 2008--2012

# Examining

-   **Matthew Warburton, University of Leeds, 8 March 2024**

-   **Nick Simonsen, University of Nottingham, 18 October 2023**

-   Danielle Beaton, University of Sheffield, 23 November, 2022

-   **Emma Henderson, Kingston University London, 17 September 2021**

-   Stanislava Naneva, University of Sheffield, 11 November 2020

-   Andrea Caso, Birkbeck, University of London, 1 July 2019**

-   **Athanasios Kokkinakis, University of York, 12 March 2019**

-   Alejandro Jimenez Rodriguez, University of Sheffield, 15 February 2019

-   **Jim Lumsden, University of Bristol, 19 October 2018**

-   Conor Heapy, University of Sheffield, 4 May 2018

-   **Joe Thompson, Simon Fraser University, Canada, 30 April 2018**

-   Jonathan Glancy, University of Sheffield, 2 December 2016

-   Christian Beyle Sandoval, University of Sheffield, 9 October 2015

-   **Maria Dagioglou, University of Birmingham, 7 April 2014**

-   Jennifer Lewis, University of Sheffield, 2012

-   David Cameron, University of Sheffield, 2011

-   Alexander Cope, University of Sheffield, 2011

-   Mike Roberts, University of Sheffield, 21st October, 2010