### 2024-07-03 news.qmd
update

---
title: "News"
editor: source
---

::: html
```{=html}
<style>
  iframe {
    width: 100%;
    aspect-ratio: 500 / 800;
    max-width: 500px; /* Optional */
  }
</style>
```
:::

## Updates from <mastodon.online/@tomstafford>

<iframe allowfullscreen sandbox="allow-top-navigation allow-scripts allow-popups allow-popups-to-escape-sandbox" src="https://mastofeed.com/apiv2/feed?userurl=https%3A%2F%2Fmastodon.online%2Fusers%2Ftomstafford&theme=light&size=100&header=false&replies=false&boosts=false"></iframe>