### 2024-07-04 done.qmd
update

---
title: "Done"
---

Recording things i've done which don't go under  [Papers](papers.html), 
[Talks](https://tomstafford.sites.sheffield.ac.uk/outputs/talks) or Newsletter: <https://tomstafford.substack.com/>. For current preoccupations, see [here](news.html).

## 2024

2024-06-21: Research Matters magazine (with Shaun Leamon): [Experiments in research funding to reduce research](https://the-sra.org.uk/common/Uploaded%20files/Research%20Matters%20Magazine/sra-research-matters-june-2024-edition.pdf) bureaucracy

2024-06-18: Early results from our project with Volkswagen Foundation: [Researchers optimistic about Distributed Peer Review](https://researchonresearch.org/dpr-volkswagen-foundation-first-results/)

2024-06-17: Hosted [AI replication game](https://sites.google.com/sheffield.ac.uk/orwg/research-practice) at University of Sheffield with the [Institute for Replication](https://i4replication.org/)

2024-05-23: Organised networking event for [University of Sheffield editors](https://sites.google.com/sheffield.ac.uk/orwg/research-practice)

2024-06-21: [2024 Showcase for Data Analysis and Visualisation MSc Students](https://tomstafford.github.io/psy6422/class-of-2024.html)

Febuary 2024: Oliver Bray shared his MSc project: Bray, O. J., & Stafford, T. (2024, February 5). [No Far-Transfer from Ten Years of Feedback on Task Length Estimation: A Case Study](https://osf.io/preprints/psyarxiv/fsr53). https://doi.org/10.31234/osf.io/fsr53.
  - see also [Ollie's blog: Can Feedback Improve Your Productivity Estimates? Data from a 10-Year Case Study
Ollie Bray](https://shinkai0202.github.io/PSY6009_Research_Project/)

2024-01-03: BBC Future article [Tetris: How a US teenager achieved the 'impossible' and what his feat tells us about human capabilities](https://www.bbc.co.uk/future/article/20240103-tetris-how-a-us-teenager-achieved-the-impossible)

## 2023

December 2023 : joined the PsyArXiv Scientific Advisory Board and approved as chair of the public engagement subcommittee. <https://osf.io/preprints/psyarxiv>

2023-11-10; Worked with Lingyu-Meng on a novel visualisation of how group discussion shifts people's opinions : <a href="Dynamics of group discussion in agreement-error space">https://lingyu-meng.github.io/convergence-valence_space/</a>

2023-11-07-: elected Chair of UKRN Institutional Leads group

Sept 2022 - Sept 2023 Coordinator for UKRN [Open Research programme](https://www.ukrn.org/open-research-programme/), [workstream 1: training](https://www.ukrn.org/ws1-training/)

2023 August made <https://tomstafford.github.io/imdb/>, showing change in crowdsourced ratings of summer movies:

<img src="https://tomstafford.github.io/imdb/plots/average_ratings_ordered.png" height=400>

2023 July Consulted on [University of Sheffield Guidelines on Generative AI](https://sheffield.libguides.com/genai/)

2023-05-23 Analysis of Open Editors data: <https://tomstafford.github.io/editors/>

<img src="https://tomstafford.github.io/editors/figs/gpa_vs_eds.png" height = 400>

2023 onwards: member of University of Sheffield Research Culture Steering Board, which provides leadership and coordination for the development of institutional research culture.

2023-05-18 Chaired UKRN/RoRI panel on [The Future of Peer Review](https://www.ukrn.org/event/future-of-peer-review-may2023/). ([Recording](https://youtu.be/BMeSh6esehE), [Slides](https://osf.io/ce3ts/))

2023-05-05 Co-author on UKRN working paper on [Open Research Training Resources and Priorities](https://osf.io/s2f6k/), ([news item](https://www.ukrn.org/2023/05/05/ukrn-first-working-paper-on-training-priorities-and-resources/))

2023-05-10 Cited in Science, Innovation and Technology Committee report on [Reproducibility and research integrity](https://publications.parliament.uk/pa/cm5803/cmselect/cmsctech/101/summary.html). Written evidence:[Commercial involvement in academic publishing is key to research reliability and should face greater public scrutiny](https://osf.io/preprints/metaarxiv/rjmvh)

2023-04-15 Chaired [Straight to the Facts: Can we save public debate? A dialogue with Will Moy](https://www.eventbrite.co.uk/e/straight-to-the-facts-can-we-save-public-debate-a-dialogue-with-will-moy-tickets-546336496167)

2023-02-19 Updated my minimal working example of RMarkdown + papaja + tenzing for APA formatted manuscript preparation, to include a rights retention statement : <https://github.com/tomstafford/rmarkdown_apa>

2023-02-06 Teaching my MSc course in Data Management and Visualisation. Materials, all CC-BY, and containing many links and resources on all things research project management, data wrangling and #dataviz  here <https://tomstafford.github.io/psy6422/>

2023-02-01 and forever: taking part in [UCU strike action](https://mastodon.online/@tomstafford/109788334770383538)

2023-01-24 Attended launch of [N8 Research Partnership stands up for researchers with new Rights Retention statement](https://www.n8research.org.uk/n8-research-partnership-rights-retention-statement/)

2023-01-17 Celebrating [University of Sheffield announcing rights retention policy](https://www.researchprofessional.com/0/rr/news/uk/universities/2023/1/University-of-Sheffield-approves-policies-to-bolster-open-research.html)

2023-01-16 Retired the [Choice Engine](https://tomstafford.github.io/choice-engine-text/) bot

2023-01-09 Graduation for Psychological Research Method with Data Science MSc [class of 2021](https://mastodon.online/@tomstafford/109659193946308768)

2023-01-06 Updated: Experiment Design Checklist <https://tomstafford.github.io/psy-checklist/>

2023-01-05 [Christmas message 2022](https://tomstafford.sites.sheffield.ac.uk/news/christmas-message-2022)

2023-01-01 [Archive of tweets 2008-2022](https://idiolect.org.uk/tweets/)

# 2021

2021 onwards: member, University of Sheffield Open Research Advsory Group, "providing strategic leadership in the ongoing development, promotion and implementation of the policies and practices supporting the University’s
commitment to Open Research"

2021 to 2024: member of University of Sheffield Research Excellence Working group, which developed, consulted on and agreed a common definition and understanding of ‘excellence’ for use alongside the institutional research strategy.

2021: [The Experiment Design Checklist](https://tomstafford.github.io/psy-checklist/)

Led development and adoption of the [University of Sheffield statement on Open Research](https://www.sheffield.ac.uk/openresearch/university-statement-open-research)

# 2020

2020: [The Letters of Pope Gregory VII - pilot project](https://figshare.com/articles/dataset/The_Letters_of_Pope_Gregory_VII_-_pilot_project/12781049) (with Charles West and George Litchfield)

2020 onwards: member, University Vice President for Research's Strategy Group (VPSG)

# 2019

[Animated Itinerary of King John,](https://sites.google.com/d/1nQb4g6jqk7M0gwaD4Bvw7QKtBg0nkQRV/p/120fMmXEkjhEwEnsgf8fx5v1XKQrf3MdZ/edit)  [Mapping The Itinerary of King Edward I](https://figshare.com/articles/dataset/Mapping_The_Itinerary_of_King_Edward_I/8948699), [Mapping Papal Letters](https://tomstafford.github.io/PapalCorrespondence/) (with Charles West)

# 2018

2018: [@ChoiceEngine - an interactive essay about the psychology of free will](https://github.com/tomstafford/choice-engine-text)

# Earlier

2017 to 2021: member, University of Sheffield Open Access Advisory Group, representing Faculty of Science, providing strategic overview of scholarly communication

2017 Organised [Mind Matters](https://tomstafford.sites.sheffield.ac.uk/sub-sites/mind-matters) 'a series of events based on various psychological research to bring knowledge to the general public and showcase Sheffield as a hub for psychology!'

2015 Data Scientist in Residence at [folksy.com](folksy.com). See [Folksy data analysis: buyer decision time](http://blog.folksy.com/2015/03/09/folksy-data-analysis-buyer-decision-time)

2012 - 2017 Wrote a column for [BBC Future](https://www.bbc.com/future/columns/neurohacks)

2012-09 I wrote some of the code in LongArm, [a robot which draws labrynthine patterns](https://www.youtube.com/watch?v=lSuvSsDrEAw), by Matthias Jones

2012: [The Cognitive Science Safari](https://sites.google.com/d/1nQb4g6jqk7M0gwaD4Bvw7QKtBg0nkQRV/p/1Q0M0b1kRUveC9w0Li7Je0A1f1E8PzzYB/edit) was a street-tour of classic social psychology experiments, in Berlin as part of the BMW Guggenheim Lab.

2011: [The Tea Taste Test](https://sites.google.com/d/1nQb4g6jqk7M0gwaD4Bvw7QKtBg0nkQRV/p/1O7uZnZPiKBL7C6JiR0DLSgGCqVw86uYX/edit) -- explaining why statistics are important to psychology, using a controversy of how to make a cup of tea.