### 2025-07-15 done.qmd
update

---
title: "Done"
---

Recording things i've done which don't go under  [Papers](papers.html), 
[Talks](https://tomstafford.sites.sheffield.ac.uk/outputs/talks) or Newsletter: <https://tomstafford.substack.com/>. For current preoccupations, see [here](news.html).

## 2025

2025-07-01 Science runs a story on our work on Distributed Peer Review: [Should grant applicants judge competitors’ proposals? Unorthodox approach gets two real-world tests](https://www.science.org/content/article/should-grant-applicants-judge-competitors-proposals-unorthodox-approach-gets-two-real) doi: 10.1126/science.zap69fh

2025-07-01 Quoted in Nature for this story on Distributed Peer Review: [How to speed up peer review: make applicants mark one another](https://www.nature.com/articles/d41586-025-02090-z). Nature 643, 313-314 (2025) doi: https://doi.org/10.1038/d41586-025-02090-z

June 29th: Helped organise and run ECR Metascientst Happy Hour, a networking event before Metascience 2025 hosted by the UKRN. Report: [How to make a career in Metascience: Advice from Experts](https://www.ukrn.org/2025/07/04/ecr-metascientist-happy-hour/)

June 16-18th: Panel DPR on DPR at International Conference on the Science of Science and Innovation ([ICSSI](https://icssi.org/)), Copenhagen

June: Stafford, T., Pinfield, S., Butters, A., & Benson Marshall, M. (2025). RoRI Insights: Applicants as reviewers - a Guide to Distributed Peer Review (Version 1). Research on Research Institute. <https://doi.org/10.6084/m9.figshare.29270534.v1>

May : external advisor and auditor for UKRI's first trial of Distributed Peer Review (read about this in their report [A Year in Metascience](https://www.gov.uk/government/publications/a-year-in-metascience-2025), published 2025-06-30)

May - June: [AFIRE Sprint on AI in grantmaking](https://researchonresearch.org/afire-ai-in-funding-sprint/)

2025-03-04: Chaired the [UKRN](https://www.ukrn.org/) Institutional Leads retreat in Cardiff. Here we are:

<img src="images/2025-03-04ukrn_ils.jpeg" height = 400>

2025-01-14: republished on the LSE Impact Blog [Do Community Notes work?](https://blogs.lse.ac.uk/impactofsocialsciences/2025/01/14/do-community-notes-work/)

## 2024

2024-12-23: Made a webscraping bot on mastodon: [https://mastodon.social/@PAXscraper](https://mastodon.social/@PAXscraper), which reports stats from PsyArXiv:

<img src="images/PAXdec2024.png" height = 400>

2024-12-19: Kate took this picture in [Elm](https://www.elmsheffield.co.uk/)

<a href="images/tom_2024-12-19.jpg"><img src="images/tom_2024-12-19small.jpg" height = 400></a>

2024-11-28/29: Represented [UKRN](https://www.ukrn.org/) at the [Swiss National Science Foundation](https://www.snf.ch/en)'s Research on Research Conference, Bern.

2024-10-23: Inaugural event of the [AFIRE Funder Forum](https://researchonresearch.org/afire-funders-forum/)

2024-10-07: Nature editorial on our project with VolkswagenStiftung: [New peer-review trial lets grant applicants evaluate each other’s proposals](https://www.nature.com/articles/d41586-024-03106-w)

2024-09: Trip to China. News story on this: [Navigating Research Culture and Funding Practices: Highlights from Shanghai](https://researchonresearch.org/shanghai-forum-tom-stafford-afire/)

2024-08-30: Article in New Scientist with Kate Dommett: [Is digital technology really swaying voters and undermining democracy?](https://www.newscientist.com/article/mg26335062-800-is-digital-technology-really-swaying-voters-and-undermining-democracy/). My [newsletter on this](https://tomstafford.substack.com/p/the-truth-about-digital-propaganda)

2024-06-21: Research Matters magazine (with Shaun Leamon): [Experiments in research funding to reduce research bureaucracy](https://the-sra.org.uk/common/Uploaded%20files/Research%20Matters%20Magazine/sra-research-matters-june-2024-edition.pdf)

2024-06-18: Early results from our project with Volkswagen Foundation: [Researchers optimistic about Distributed Peer Review](https://researchonresearch.org/dpr-volkswagen-foundation-first-results/)

2024-06-17: Hosted [AI replication game](https://sites.google.com/sheffield.ac.uk/orwg/research-practice) at University of Sheffield with the [Institute for Replication](https://i4replication.org/)

2024-05-23: Organised networking event for [University of Sheffield editors](https://sites.google.com/sheffield.ac.uk/orwg/research-practice)

2024-06-21: [2024 Showcase for Data Analysis and Visualisation MSc Students](https://tomstafford.github.io/psy6422/class-of-2024.html)

Febuary 2024: Oliver Bray shared his MSc project: Bray, O. J., & Stafford, T. (2024, February 5). [No Far-Transfer from Ten Years of Feedback on Task Length Estimation: A Case Study](https://osf.io/preprints/psyarxiv/fsr53). https://doi.org/10.31234/osf.io/fsr53.
  - see also [Ollie's blog: Can Feedback Improve Your Productivity Estimates? Data from a 10-Year Case Study
Ollie Bray](https://shinkai0202.github.io/PSY6009_Research_Project/)

2024-01-03: BBC Future article [Tetris: How a US teenager achieved the 'impossible' and what his feat tells us about human capabilities](https://www.bbc.co.uk/future/article/20240103-tetris-how-a-us-teenager-achieved-the-impossible)

## 2023

December 2023 : joined the PsyArXiv Scientific Advisory Board and approved as chair of the public engagement subcommittee. <https://osf.io/preprints/psyarxiv>

2023-11-10; Worked with Lingyu-Meng on a novel visualisation of how group discussion shifts people's opinions : <a href="Dynamics of group discussion in agreement-error space">https://lingyu-meng.github.io/convergence-valence_space/</a>

2023-11-07-: elected Chair of UKRN Institutional Leads group

Sept 2022 - Sept 2023 Coordinator for UKRN [Open Research programme](https://www.ukrn.org/open-research-programme/), [workstream 1: training](https://www.ukrn.org/ws1-training/)

2023 August made <https://tomstafford.github.io/imdb/>, showing change in crowdsourced ratings of summer movies:

<img src="https://tomstafford.github.io/imdb/plots/average_ratings_ordered.png" height=400>

2023 July Consulted on [University of Sheffield Guidelines on Generative AI](https://sheffield.libguides.com/genai/)

2023-05-23 Analysis of Open Editors data: <https://tomstafford.github.io/editors/>

<img src="https://tomstafford.github.io/editors/figs/gpa_vs_eds.png" height = 400>

2023 onwards: member of University of Sheffield Research Culture Steering Board, which provides leadership and coordination for the development of institutional research culture.

2023-05-18 Chaired UKRN/RoRI panel on [The Future of Peer Review](https://www.ukrn.org/event/future-of-peer-review-may2023/). ([Recording](https://youtu.be/BMeSh6esehE) and DOI:[10.52843/cassyni.50wy02](https://doi.org/10.52843/cassyni.50wy02), [Slides](https://osf.io/ce3ts/))

2023-05-05 Co-author on UKRN working paper on [Open Research Training Resources and Priorities](https://osf.io/s2f6k/), ([news item](https://www.ukrn.org/2023/05/05/ukrn-first-working-paper-on-training-priorities-and-resources/))

2023-05-10 Cited in Science, Innovation and Technology Committee report on [Reproducibility and research integrity](https://publications.parliament.uk/pa/cm5803/cmselect/cmsctech/101/summary.html). Written evidence:[Commercial involvement in academic publishing is key to research reliability and should face greater public scrutiny](https://osf.io/preprints/metaarxiv/rjmvh)

2023-04-15 Chaired [Straight to the Facts: Can we save public debate? A dialogue with Will Moy](https://www.eventbrite.co.uk/e/straight-to-the-facts-can-we-save-public-debate-a-dialogue-with-will-moy-tickets-546336496167)

2023-02-19 Updated my minimal working example of RMarkdown + papaja + tenzing for APA formatted manuscript preparation, to include a rights retention statement : <https://github.com/tomstafford/rmarkdown_apa>

2023-02-06 Teaching my MSc course in Data Management and Visualisation. Materials, all CC-BY, and containing many links and resources on all things research project management, data wrangling and #dataviz  here <https://tomstafford.github.io/psy6422/>

2023-02-01 and forever: taking part in [UCU strike action](https://mastodon.online/@tomstafford/109788334770383538)

2023-01-24 Attended launch of [N8 Research Partnership stands up for researchers with new Rights Retention statement](https://www.n8research.org.uk/n8-research-partnership-rights-retention-statement/)

2023-01-17 Celebrating [University of Sheffield announcing rights retention policy](https://www.researchprofessional.com/0/rr/news/uk/universities/2023/1/University-of-Sheffield-approves-policies-to-bolster-open-research.html)

2023-01-16 Retired the [Choice Engine](https://tomstafford.github.io/choice-engine-text/) bot

2023-01-09 Graduation for Psychological Research Method with Data Science MSc [class of 2021](https://mastodon.online/@tomstafford/109659193946308768)

2023-01-06 Updated: Experiment Design Checklist <https://tomstafford.github.io/psy-checklist/>

2023-01-05 [Christmas message 2022](https://tomstafford.sites.sheffield.ac.uk/news/christmas-message-2022)

2023-01-01 [Archive of tweets 2008-2022](https://idiolect.org.uk/tweets/)

# 2021

2021 onwards: member, University of Sheffield Open Research Advsory Group, "providing strategic leadership in the ongoing development, promotion and implementation of the policies and practices supporting the University’s
commitment to Open Research"

2021 to 2024: member of University of Sheffield Research Excellence Working group, which developed, consulted on and agreed a common definition and understanding of ‘excellence’ for use alongside the institutional research strategy.

2021: [The Experiment Design Checklist](https://tomstafford.github.io/psy-checklist/)

Led development and adoption of the [University of Sheffield statement on Open Research](https://www.sheffield.ac.uk/openresearch/university-statement-open-research)

# 2020

2020: [The Letters of Pope Gregory VII - pilot project](https://figshare.com/articles/dataset/The_Letters_of_Pope_Gregory_VII_-_pilot_project/12781049) (with Charles West and George Litchfield)

2020 onwards: member, University Vice President for Research's Strategy Group (VPSG)

# 2019

[Animated Itinerary of King John,](https://sites.google.com/d/1nQb4g6jqk7M0gwaD4Bvw7QKtBg0nkQRV/p/120fMmXEkjhEwEnsgf8fx5v1XKQrf3MdZ/edit)  [Mapping The Itinerary of King Edward I](https://figshare.com/articles/dataset/Mapping_The_Itinerary_of_King_Edward_I/8948699), [Mapping Papal Letters](https://tomstafford.github.io/PapalCorrespondence/) (with Charles West)

# 2018

2018: [@ChoiceEngine - an interactive essay about the psychology of free will](https://github.com/tomstafford/choice-engine-text)

# Earlier

2017 to 2021: member, University of Sheffield Open Access Advisory Group, representing Faculty of Science, providing strategic overview of scholarly communication

2017 Organised [Mind Matters](https://tomstafford.sites.sheffield.ac.uk/sub-sites/mind-matters) 'a series of events based on various psychological research to bring knowledge to the general public and showcase Sheffield as a hub for psychology!'

2015 Data Scientist in Residence at [folksy.com](folksy.com). See [Folksy data analysis: buyer decision time](http://blog.folksy.com/2015/03/09/folksy-data-analysis-buyer-decision-time)

2012 - 2017 Wrote a column for [BBC Future](https://www.bbc.com/future/columns/neurohacks)

2012-09 I wrote some of the code in LongArm, [a robot which draws labrynthine patterns](https://www.youtube.com/watch?v=lSuvSsDrEAw), by Matthias Jones

2012 September: [The Cognitive Science Safari](done/2012-09-06.html) was a street-tour of classic social psychology experiments, in Berlin as part of the BMW Guggenheim Lab. I was on the cover of the Wall Street Journal.

2012-06-06: [Fundamentals of learning: the exploration-exploitation trade-off](done/2012-06-12.html)

2012-06-04: [New paper: A novel task for the investigation of action acquisition](done/2012-06-12.html)

2011: [The Tea Taste Test](https://sites.google.com/d/1nQb4g6jqk7M0gwaD4Bvw7QKtBg0nkQRV/p/1O7uZnZPiKBL7C6JiR0DLSgGCqVw86uYX/edit) -- explaining why statistics are important to psychology, using a controversy of how to make a cup of tea.