### 2025-08-01 index.qmd
update

---
title: "Tom Stafford"
format: html
page-layout: full
image: /tom_2020.jpg
---
::: {.column-page-inset}

![](images/tom_2020.jpg){width="50%"}

As Professor of Cognitive Science, I teach and research in the [School of Psychology](https://www.sheffield.ac.uk/psychology) at the University of Sheffield. My work applies insights from neuroscience, experimental design and data science to the topics of learning, decision making, reasoning and bias.

::: {.callout-tip title="News!"}
From 1st September 2025 - 31st August 2026 I am on leave from the University of Sheffield. During this time I will be on secondment to the [Research on Research Institute](https://researchonresearch.org/), as well as pursing writing projects. Subscribe for updates on [tomstafford.substack.com/](https://tomstafford.substack.com/).
:::

As Senior Fellow at the [Research on Research Institute](https://researchonresearch.org/), I lead the programme [AFIRE: Accelerator For Innovation &amp; Research Funding Experimentation](https://researchonresearch.org/project/a-f-i-r-e/). We seek to catalyse the generation and use of high quality evidence on research processes ("Metaresearch").

I am on the [PsyArXiv](https://osf.io/preprints/psyarxiv) Scientific Advisory Board as chair of the Public Engagement Subcommittee.

Previously, I was Director for the MSc programmes in Psychological Research Methods, and set up the MSc in Psychological Research Methods with Data Science. You can read more about the flagshiop module I designed and taught here: [PSY6422 Data Analysis and Visualisation](https://tomstafford.github.io/psy6422/).

From 2020 to 2025 I held the Research Practice Lead for the University of Sheffield, an institution-wide role with a remit to encourage practices which support robust and trusted research. As Research Practice Lead, I represented the University of Sheffield in the [UK Reproducibility Network](https://www.ukrn.org/), for which I chaired the national [UKRN Institutional Leads group](https://www.ukrn.org/institutional-leads/) 2023 - 2025.

For non-specialist audiences, I have written a [best-selling book](https://mindhacks.com/book/), an [award winning blog](https://mindhacks.com/) and columns for venues including BBC, New Scientist, The Guardian and The Conversation. I have also written some mediocre selling books, some widely ignored blogs and tweeted jokes which I thought were hilarious at the time but which received precisely zero likes and zero retweets.

:::

These pages are intended to replace [tomstafford.sites.sheffield.ac.uk](https://tomstafford.sites.sheffield.ac.uk/), and will reflect my professional activities.

My non-professional activities are recorded on my personal blog at <https://idiolect.org.uk/notes/>

Or you can find traces on these other platforms:

{{< pagebreak >}}

::: {.column-page-inset .text-center}
```{=html}
<a class="btn btn-outline-dark", href="https://tomstafford.substack.com" target="_blank" rel="noopener noreferrer">
    <i class="bi bi-substack" role='img' aria-label='Substack'></i>
</a>

<a class="btn btn-outline-dark", href="https://mastodon.online/@tomstafford" target="_blank" rel="noopener noreferrer">
    <i class="bi bi-mastodon" role='img' aria-label='Mastodon'></i>
</a>

<a class="btn btn-outline-dark", href="https://github.com/tomstafford" target="_blank" rel="noopener noreferrer">
    <i class="bi bi-github" role='img' aria-label='Github'></i>
</a>

<a class="btn btn-outline-dark", href="https://www.linkedin.com/in/tom-stafford-81a737262/" target="_blank" rel="noopener noreferrer">
    <i class="bi bi-linkedin" role='img' aria-label='LinkedIn'></i>
</a>

<a class="btn btn-outline-dark", href="https://scholar.google.co.uk/citations?user=bMEna2UAAAAJ&hl=en&oi=ao" target="_blank" rel="noopener noreferrer">
    <i class="ai ai-google-scholar" role='img' aria-label='Google Scholar'></i>
</a>

<a class="btn btn-outline-dark", href="mailto:t.stafford@sheffield.ac.uk" target="_blank" rel="noopener noreferrer">
    <i class="ai bi-envelope" role='img' aria-label='email'></i>
</a>
```
:::

{{< pagebreak >}}

## More

* [Research](research.html)
* [Papers](papers.html)
* [Writing](writing.html)
* [Talks](talks.html)
* [Teaching](teaching.html)
* [Supervision](supervision.html)
* [Done](done.html)
* [News](news.html)

## Colophon

Website made with [Quarto](https://quarto.org), hosted on [github](https://github.com/tomstafford/tomstafford.github.io). Site style & structure from [Drew Dimmery](https://ddimmery.com/posts/quarto-website/).